---
title: "Working with Adjustment Specifications"
author: "Chris Hansen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Adjustment Specifications}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

<style type="text/css">
  .section.level1 h1 {
    font-size: 25px;
    line-height: 30px;
    margin-top: 20px;
    border-bottom: 6px solid #f7f7f7;
  }

.figure img{
  max-width: 80%;
}

.plotly.html-widget{
  width: 700px !important;
  padding-bottom: 10px;
}
</style>
  
```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE,
                      message = FALSE, comment = NA, 
                      fig.align = "center", 
                      fig.height = 4, fig.width = 5)
library(sadj)
```

<!--
# `x13binary`

The `sadj` package has several dependencies, the most significant being the
[`x13binary`](https://cran.r-project.org/web/packages/x13binary/index.html)
package which provides the `X13-ARIMA-SEATS` binary.  

Ordinarily, installing the CRAN version of `x13binary` will cause a prebuilt
binary to be downloaded, but the Stats NZ firewall prevents this happening.
Instead, we have produced a modified vesion of the `x13binary` package which
includes a version of the `X13-ARIMA-SEATS` binary built from sources found on
the [U.S. Census Bureau website](https://www.census.gov/ts/x13as/unix/x13ashtmlsrc_V1.1_B39.tar.gz).
Users can install this to their own personal library via the following command
(assuming `svn` is working correctly):

```r
devtools::install_svn("https://WPRDSVN01.stats.govt.nz/svn/RSTUDIO/StatsMeth/packages/x13binary")
```

Ultimately, a properly functioning version of this package will be available in
the usual shared libraries, and users will not be required to install it
themselves.
-->
<!--
# `seasonal`

The first version of the `sadj` package was written a number of years ago and
used for monitoring of the Household Labour Force Survey.  Since then, a package
called [`seasonal`](https://cran.r-project.org/web/packages/seasonal/index.html) 
has been released to CRAN which is somewhat similar.

However, as run by Stats NZ, the `seasonal` package has some limitations.  The
most significant is the inability to group series together in a single
adjustment via an `mta` file, including the adjustment of composites.  

The `seasonal` package is well constructed, though, and could still prove useful
as a tool for analysing unadjusted time series data, series by series.  In
particular, the package comes with a `shiny` interface, a demo of which is
available [here](http://www.seasonal.website/).
-->
<!--
# A Note on External Data Files

Those uses who have used the `X13-ARIMA-SEATS` program directly will know that
it is possible to provide the raw series to be adjusted via either a `data` or
`file` parameter in the `series` specification.  The former is a delimited
list of the input series values, while the latter is the name of a file which
holds the raw data.  

For the implementation provided here, the user is not required to provide
either.  An `X13Series` object _is_ the raw data, and it is inserted into the
specification via the `data` parameter mentioned at the time of adjustment.  In
fact, if either of `data` or `series` (along with the related argument
`format`, if present) parameters are found inside the `series` specification
they are ignored.
--> 

There are two general methods for applying specifications for an adjustment:

* provide a complete specification when creating an adjustable object
* modify the specification of an existing adjustable object

We will show examples of both in this document.


## Background

The `sadj` package provides a wrapper for the X13-ARIMA-SEATS program.  This is
a command line program, the minimal input requirement of which is a so-called
specification file.  A specification file is a text file with extension `spc`
consisting of blocks of adjustement parameters as follows:

```as.is
specfication1{
  parameter1=value1
  parameter2=value2
  ...
}

specification2{
  parameter1=value1
  parameter2=value2
  ...
}

...
```

For example, a modified version of an example specification which is usually
provided with the X13-ARIMA-SEATS program looks as follows:

```as.is
series{
  start=1949.01
  data=(
	112 118 132 129 121 135 148 148 136 119 104 118
	115 126 141 135 125 149 170 170 158 133 114 140
	145 150 178 163 172 178 199 199 184 162 146 166
	171 180 193 181 183 218 230 242 209 191 172 194
	196 196 236 235 229 243 264 272 237 211 180 201
	204 188 235 227 234 264 302 293 259 229 203 229
	242 233 267 269 270 315 364 347 312 274 237 278
	284 277 317 313 318 374 413 405 355 306 271 306
	315 301 356 348 355 422 465 467 404 347 305 336
	340 318 362 348 363 435 491 505 404 359 310 337
	360 342 406 396 420 472 548 559 463 407 362 405
	417 391 419 461 472 535 622 606 508 461 390 432)
}

x11{
  save=(d10 d11 d12)
}
```

Many required parameters have default values, so don't need to be speficied
explicitly.  In this case, all that is supplied is a minimal `series` block, and
a minimal `x11` block. The `series` block is used to provide the raw series;
and the `x11` block means that the X11 algorithm will be used, rather than the
SEATS method of adjustment, and series named `d10`, `d11`, and `d12` will be
saved in external files (being the seasonal factors, the seasonally adjusted
series, and the trend series, respectively).  Assuming the specification was
saved as `airline.spc`, the adjustment would be run from the command line via
something like:

```bash
x13ashtml airline
```

This is the same process used by the `sadj` package, though the details are
hidden from users, and there are several minor modifications which will be
noted in this document.  Importantly, specifications are held in R lists, and
only briefly converted to physical specification files at the point of
adjustment.  

Convenience methods are provided for importing existing specification files into
R, and will be covered below.  It will also be possible to work with
specifications in JSON format, which might be a preferred format for some,
particularly if the goal is to build a web service via OpenCPU, say.


## Default specifications

The simplest possible set of calls that yield an adjusted time series are as
follows:

```{r simple}
ap <- X13Series(AirPassengers)
ap.res <- adjust(ap)
```

In an earlier example, we saw a simple adjustment without any specifications
being explicitly provided.  Actually, a default specification was provided,
and the original call to `X13Series` could have been written in the following
equivalent way:

```r
ap <- X13Series(AirPassengers, type = "x11")
```

If the SEATS method is preferred over X11, then a default SEATS adjustment is
also readily available:

```{r seats01}
ap <- X13Series(AirPassengers, type = "seats")
getSpecList(ap)
```

## Creating objects from scratch

## Modifying existing objects

An easy method is to decide whether one wants an X11 or SEATS adjustment,
create an object with the appropriate default, and then modify the result.






